
DSA306 Group 1 Team 5
By: Li Song Xi (01379236), Cavin Ang Kie King (01353493), Joanna Jee Xu Hui (01394470)

# Getting MRT, Schools, and Mall distance
```{python}
import pandas as pd
import numpy as np
from bs4 import BeautifulSoup
import requests
import lxml
import googlemaps
from urllib.parse import urlencode

api_key = "AIzaSyAFkE53oqBILQtY3adBs9eu6fSXiAyIDlI"

#Gmaps Function
def extract_lat_lng(name):
  endpoint = f"https://maps.googleapis.com/maps/api/geocode/json"
  params = {'address': name, 'key':api_key}
  url_params = urlencode(params)
  url = f"{endpoint}?{url_params}"
  r = requests.get(url)
  if r.status_code not in range(200, 299):
    return {}
  latlng = {}
  try:
    latlng = r.json()['results'][0]['geometry']['location']
  except:
    pass
  return latlng['lat'], latlng['lng']


#Pri Sch Distance

pri_sch_url = "https://en.wikipedia.org/wiki/List_of_primary_schools_in_Singapore"
pri_html = requests.get(pri_sch_url, allow_redirects = True)
pri_soup = BeautifulSoup(pri_html.text, 'html.parser')
pri_table = pri_soup.find('table', attrs = {'class': "wikitable sortable"})
pri_sch = pd.read_html(str(pri_table))
pri_sch = pd.concat(pri_sch)

pri_sch['Name'] = pri_sch['Name']+", Singapore"
# pri_names = df['Name']
pri_sch['latlng'] = pri_sch['Name'].apply(extract_lat_lng)

lat_list = []
lng_list = []
for lat, lng in pri_sch['latlng']:
  lat_list.append(lat)
  lng_list.append(lng)
pri_sch['lat'] = lat_list
pri_sch['lng'] = lng_list

pri_sch.to_csv("raw_data/pri_sch_modified.csv")


#Sec sch Distance

sec_sch_url = "https://en.wikipedia.org/wiki/List_of_secondary_schools_in_Singapore"
sec_html = requests.get(sec_sch_url, allow_redirects = True)
sec_soup = BeautifulSoup(sec_html.text, 'html.parser')
sec_table = sec_soup.find('table', attrs = {'class': "wikitable sortable"})
sec_sch = pd.read_html(str(sec_table))
sec_sch = pd.concat(sec_sch)

sec_sch['Name'] = sec_sch['Name']+", Singapore"
sec_sch['latlng'] = sec_sch['Name'].apply(extract_lat_lng)

lat_list = []
lng_list = []
for lat, lng in sec_sch['latlng']:
  lat_list.append(lat)
  lng_list.append(lng)
sec_sch['lat'] = lat_list
sec_sch['lng'] = lng_list

sec_sch.to_csv("raw_data/sec_sch_modified.csv")
```

```{python}
from googlemaps import *
import pandas as pd 

# API
gmaps = Client(key = 'AIzaSyBXmbqQFtcEily7FRof-jHUkDv1nYWKIXw')

addresses = pd.read_csv("raw_data/shoppingmall.csv")
addresses.head()

# Creating colummns
addresses['lng'] = ""
addresses['lat'] = ""

for x in range(len(addresses)):
    try:
        geocode_result = gmaps.geocode(addresses['address'][x])
        addresses['lat'][x] = geocode_result[0]['geometry']['location'] ['lat']
        addresses['lng'][x] = geocode_result[0]['geometry']['location']['lng']
    except IndexError:
        print("Address was wrong...")
    except Exception as e:
        print("Unexpected error occurred.", e )
addresses.head()

# Save to csv
addresses.to_csv('raw_data/shopping_malls_coords.csv')
```

```{r}
library(readr)
mrt = read.csv('raw_data/Mrt.csv')
colnames(mrt)
# MRT stations names
mrt$full_name = paste(mrt$ï..mrt_station_english, mrt$mrt, mrt$stn_code)
mrt$full_name
mrt_list = toString(mrt$full_name)
mrt_list

# Geocode to obtain Longs & Lats
library(ggmap)
library(tmaptools)
library(RCurl)
library(jsonlite)
library(tidyverse)
library(leaflet)
library(googleway)
register_google(key = 'AIzaSyD6MHvuFCm0n9Wot8VRPuJwuVF7AkkHwas')
mrt_list_new = unlist(strsplit(mrt_list,","))
mrt_df <- data.frame(MRT = mrt_list_new, stringsAsFactors = FALSE)
mrt_ggmap <- geocode(location = mrt_list_new, output = "more", source = "google")
mrt_ggmap <- cbind(mrt_df, mrt_ggmap)
long_lat = mrt_ggmap[, 1:3]
long_lat |> rename('Name' = 'MRT', 'lng' = 'lon')


long_lat1 = long_lat |> rename('Name' = 'MRT', 'lng' = 'lon')
which(is.na(long_lat1$lng))
long_lat1$Name[63]
long_lat1$lng[63] = 103.8215381  
long_lat1$lat[63] = 1.2650372
long_lat1$Name[149]
long_lat1$lng[149] = 103.7666228  
long_lat1$lat[149] = 1.377783623
#manually enter long-lat data as previous function could not find

# Write to csv to update NA values with proper coords
write.csv(long_lat1,"raw_data/mrt_lng_lat.csv", row.names = FALSE)
```


```{python}
import pandas as pd
import haversine as hs


def add_to_main(main_df, dist_df, ver):
  """Uses min_dist function to find each addresses closest location within dist_df to each address in main_df"""
  sch_list = []
  dist_list = []
  for index, row in main_df.iterrows():
    x = min_dist(row['latitude'], row['longitude'], dist_df = dist_df)
    sch_list.append(x[0])
    dist_list.append(x[1])
  main_df[f'closest_{ver}'] = sch_list
  main_df[f'{ver}_dist'] = dist_list

def min_dist(lat, lng, dist_df):
  """Uses hs.haversine to calculate distance from each location based on lat/lng and returns closeset address and its distance"""
  dist = []
  for index, row in dist_df.iterrows():
    dist.append(hs.haversine((lat, lng), (row['lat'], row['lng'])))
  pos = dist.index(min(dist))
  return dist_df.iloc[pos]['Name'], min(dist)
```

```{python}  
pri_sch_df = pd.read_csv("raw_data/pri_sch_modified.csv")[['Name', 'lat', 'lng']]
sec_sch_df = pd.read_csv("raw_data/sec_sch_modified.csv")[['Name', 'lat', 'lng']]
mall_df = pd.read_csv("raw_data/shopping_malls_coords.csv")[['Name', 'lat', 'lng']]
mrt_df = pd.read_csv("raw_data/mrt_lng_lat.csv")[['Name', 'lat', 'lng']]

mall_df.head()

hdb_df_full = pd.read_csv("raw_data/ALL Prices 1990-2021 mar.csv")
u_loc_full = hdb_df_full.drop_duplicates(['address'])[['address', 'latitude', 'longitude']]

```

Adding all to main data frame and saving as csv
```{python}
add_to_main(u_loc_full, dist_df = pri_sch_df, ver = "pri")
add_to_main(u_loc_full, dist_df = sec_sch_df, ver = "sec")
add_to_main(u_loc_full, dist_df = mall_df, ver = "mall")
add_to_main(u_loc_full, dist_df = mrt_df, ver = "mrt")

u_loc_full.head()
u_loc_full.to_csv('raw_data/distance_to_all.csv')
```

Now we have the distance for all locations!


#Cleaning with the data!!!

```{r}
library(sparklyr)
library(dplyr)

config <- spark_config()
config$`sparklyr.shell.driver-memory` <- "6G"
config$`sparklyr.shell.executor-memory` <- "6G"
sc <- spark_connect(master = "local", config = config, version = "3.3.0")

# sc <- spark_connect(master = "local", version = "3.3.0")
hdb <- spark_read_csv(sc, name = 'hdb_data', path = "raw_data/ALL Prices 1990-2021 mar.csv", header =TRUE, delimiter = ',', memory = FALSE)
distance <- spark_read_csv(sc, name = 'distance_data', path = "raw_data/distance_to_all.csv", header =TRUE, delimiter = ',', memory = FALSE)

library(DBI)
library(dbplot)
library(ggplot2)
library(gridExtra)

merged_data <- inner_join(hdb, distance, by = 'address') 
```


Rows that are removed from the data-set
```{r}
truncated_data <- select(merged_data, - c('month', 'town_dummy', 'block', 'address', 'street_name', 'latitude_x', 'longitude_y', 'storey_range', 18:28, '_c0', 'latitude_y', 'longitude_x', 'closest_mrt', 'closest_mall', 'closest_pri', 'closest_sec', 'flat_model' ))

# EDA
# Check for NAs
hdb_tbl <- copy_to(sc, truncated_data, overwrite = TRUE)
sum(is.na(hdb_tbl))
data_cleaned <- hdb_tbl #|> mutate(flat_type = as.numeric(substr(flat_type, 1,2)))
head(data_cleaned)

columns <- colnames(data_cleaned)
columns

for (i in columns){
  data_cleaned[[i]] = as.double(data_cleaned[[i]])
}
data <- as.data.frame(data_cleaned)
data <- data |>
  filter(year>=2017)
data <- sdf_copy_to(sc, data, overwrite = TRUE)
spark_write_parquet(data, path = "data/combined_data.parquet")
data <- spark_read_parquet(sc, path = "data/combined_data.parquet")
```
data cleaned!

##EDA
#Correlation plot
```{r}
library(corrr)
data_cor  <- select(data, -c(flat_type, town)) 
data_cor
#remove flat type from corelation plot as its categorical
# cor(data_cor, use = "complete.obs")
data_cor |> 
  correlate(use = "pairwise.complete.obs", method = "pearson") |> 
  shave(upper = TRUE) |> 
  rplot(colours = c("indianred2", "black", "skyblue1"))
```

#Checking spread of each variable
```{r}
library(ggplot2)
library(gridExtra)
q <- c(0.25, 0.5, 0.75)
      
# Resale
ggplot(data = data, aes(x = resale_price, y = 1)) +
geom_violin(draw_quantiles = q)
# positive skew for resale price -> log
violin_resale_price <- ggplot(data = data, aes(x = log(resale_price), y = 1)) +
  geom_violin(draw_quantiles = q)
violin_resale_price
      
# area_sqm
violin_area_sqm <- ggplot(data = data, aes(x = area_sqm, y = 1)) +
  geom_violin(draw_quantiles = q)
violin_area_sqm

# lease_rem
violin_lease_rem <- ggplot(data = data, aes(x = lease_rem, y = 1)) +
  geom_violin(draw_quantiles = q) 
violin_lease_rem
#most people sell when theres more lease remaining may want to take out lease_start
      
# pri_dist
ggplot(data = data, aes(x = pri_dist, y = 1)) +
  geom_violin(draw_quantiles = q)
# positive skew for pri_dist -> log
violin_pri_dist <- ggplot(data = data, aes(x = log(pri_dist), y = 1)) +
  geom_violin(draw_quantiles = q) 
violin_pri_dist

# sec_dist
ggplot(data = data, aes(x = sec_dist, y = 1)) +
  geom_violin(draw_quantiles = q)
# positive skew for sec_dist -> log
violin_sec_dist <- ggplot(data = data, aes(x = log(sec_dist), y = 1)) +
  geom_violin(draw_quantiles = q) 
violin_sec_dist 
      
# mall_dist
ggplot(data = data, aes(x = mall_dist, y = 1)) +
  geom_violin(draw_quantiles = q)
# positive skew for mall_dist -> sqrt
violin_mall_dist <- ggplot(data = data, aes(x = sqrt(mall_dist), y = 1)) +
  geom_violin(draw_quantiles = q)
violin_mall_dist 
      
      # mrt_dist
ggplot(data = data, aes(x = mrt_dist, y = 1)) +
  geom_violin(draw_quantiles = q)
# positive skew for mrt_dist -> log
violin_mrt_dist <- ggplot(data = data, aes(x = log(mrt_dist), y = 1)) +
  geom_violin(draw_quantiles = q)
violin_mrt_dist
```

#Ploting each variable against resale_price
```{r}
library(ggplot2)
# 1. year 
#ggplot(data=data, aes(x=year, y=resale_price)) + geom_point() # Use raster plot inside
year <- data |> 
  group_by(year) |> 
  summarise(avg_resale_price = mean(resale_price))
ggplot(data=year, aes(x=year, y=avg_resale_price)) + geom_line()
# Positive relationship between year and resale price
    
# 2. 
no_rooms <- table(data$flat_type)
# freq_rooms <- as.data.frame(freq)
# freq_rooms
rooms <- data |> group_by(flat_type) |> summarise(avg_resale_price = mean(resale_price))
ggplot(data=rooms, aes(x=flat_type, y=avg_resale_price)) + geom_point()
# Strong Positive relationship between year and resale price

# 3.
storey <- data |> group_by(storey) |> summarise(avg_resale_price = mean(resale_price))
ggplot(data=storey, aes(x=storey, y=avg_resale_price)) + geom_point()
# Strong Positive relationship between storey and resale price
      
# 4.
lease_rem <- data |> group_by(lease_rem) |> summarise(avg_resale_price = mean(resale_price))
ggplot(data=lease_rem, aes(x=lease_rem, y=avg_resale_price)) + geom_point()
# No clear relationship between lease remaining and resale price
      
# 5. 
pri_dist <- data |> group_by(pri_dist) |> summarise(avg_resale_price = mean(resale_price)) |> filter(pri_dist<2)
ggplot(data=pri_dist, aes(x=pri_dist, y=avg_resale_price)) + geom_point()
# No clear relationship between pri sch distance and resale price
      
# 6.
sec_dist <- data |> group_by(sec_dist) |> summarise(avg_resale_price = mean(resale_price)) |> filter(sec_dist<2)
ggplot(data=sec_dist, aes(x=sec_dist, y=avg_resale_price)) + geom_point()
# No clear relationship between sec sch distance and resale price 
      
# 7.
mall_dist <- data |> group_by(mall_dist) |> summarise(avg_resale_price = mean(resale_price)) |> filter(mall_dist<2)
ggplot(data=mall_dist, aes(x=mall_dist, y=avg_resale_price)) + geom_point()
# Weak positive relationship between sec sch distance and resale price on prices > 500,000
      
# 8.
mrt_dist <- data |> group_by(mrt_dist) |> summarise(avg_resale_price = mean(resale_price)) |> filter(mrt_dist<2)
ggplot(data=mrt_dist, aes(x=mrt_dist, y=avg_resale_price)) + geom_point()
# Weak positive relationship between mrt distance and resale price on prices > 500,000
```



#Checking significance of dummy variables
```{r}
CI <- 1.96 #at a 95% confidence leve

dummy_town <- data |>
  group_by(town) |>
  summarise(count = n(), avg_resale_price = mean(resale_price), sd = sd(resale_price)) |>
  mutate(se = sd/sqrt(count)) |>
  collect()

dummy_town |>
  ggplot(aes(x = town, y = avg_resale_price)) +
  geom_point(size = 2) +
  geom_errorbar(
    aes(ymin = avg_resale_price-CI*se, ymax = avg_resale_price+CI*se),
    width = 0.1
  ) + geom_hline(
    yintercept = mean(dummy_town$avg_resale_price),
    linetype= "dashed"
  )


dummy_flat <- data |>
  group_by(flat_type) |>
  summarise(count = n(), avg_resale_price = mean(resale_price), sd = sd(resale_price)) |>
  mutate(se = sd/sqrt(count)) |>
  collect()

dummy_flat |>
  ggplot(aes(x = flat_type, y = avg_resale_price)) +
  geom_point(size = 2) +
  geom_errorbar(
    aes(ymin = avg_resale_price-CI*se, ymax = avg_resale_price+CI*se),
    width = 0.1
  ) + geom_hline(
    yintercept = mean(dummy_flat$avg_resale_price),
    linetype= "dashed"
  )

```
#Testing Different Models
##Spliting the Data
```{r}
data_split <- data |>
  sdf_random_split(training = 0.8, test = 0.2, seed = 1337)

split_train <- data_split$training
split_test <- data_split$test
```

```{r}

fit1 <- split_train |> ml_linear_regression(formula = resale_price ~ .)
summary(fit1)
fit1_sum <- ml_evaluate(fit1, dataset = split_test)
fit1_sum$root_mean_squared_error
fit1_sum$r2adj
# R squared value of 0.8634
```
# Transforming then Spliting 
Previously we had seen that some colums were positively skewed. Here we try to transform the data before putting it into the model
```{r}
split_trans_train <- split_train |>
  mutate(resale_price_lg = log(resale_price),
         pri_dist_lg = log(pri_dist),
         sec_dist_lg = log(sec_dist),
         mall_dist_sqrt = sqrt(mall_dist),
         mrt_dist_lg = log(mrt_dist)) |>
  select(-c(resale_price, pri_dist, sec_dist, mall_dist, mrt_dist))
split_trans_test <- split_test |>
  mutate(resale_price_lg = log(resale_price),
         pri_dist_lg = log(pri_dist),
         sec_dist_lg = log(sec_dist),
         mall_dist_sqrt = sqrt(mall_dist),
         mrt_dist_lg = log(mrt_dist)) |>
  select(-c(resale_price, pri_dist, sec_dist, mall_dist, mrt_dist))
```

fit2
```{r}
fit2 <- split_trans_train |> ml_linear_regression(formula = resale_price_lg~.)
summary(fit2)

fit2_sum <- ml_evaluate(fit2, split_trans_test)
fit2_sum$root_mean_squared_error
fit2_sum$r2adj
# R squared value of 0.893
```

fit3
```{r}
fit3 <- split_trans_train |> ml_linear_regression(formula = resale_price_lg~ town + year + flat_type + storey + area_sqm + lease_start + lease_rem+sec_dist_lg + pri_dist_lg + mall_dist_sqrt + mrt_dist_lg + flat_type*area_sqm + lease_start*lease_rem)
summary(fit3)

fit3_sum <- ml_evaluate(fit3, split_trans_test)
fit3_sum$root_mean_squared_error
fit3_sum$r2adj
# R squared value of 0.8967 after including interaction terms between lease_start, lease_rem and area_sqm, flat_type
```

#Pipeline
```{r}
sql <- 
  "
  SELECT 
    year, storey, area_sqm, lease_start, lease_rem, LOG(resale_price) as resale_price_lg, LOG(pri_dist) as pri_dist_lg, LOG(sec_dist) as sec_dist_lg, SQRT(mall_dist) as mall_dist_sqrt, LOG(mrt_dist) as mrt_dist_lg, town, flat_type
  FROM __THIS__
  "

pipeline <- ml_pipeline(sc)|>
  ft_sql_transformer(statement = sql) |>
  ft_r_formula(formula = resale_price_lg ~ town + year + flat_type + storey + area_sqm + lease_start + lease_rem + sec_dist_lg + pri_dist_lg + mall_dist_sqrt + mrt_dist_lg + flat_type*area_sqm + lease_start*lease_rem) |>
  ml_linear_regression()
# pipeline
# split_trans_train
  
pipeline_model  <- ml_fit(pipeline, split_train)
cv <- ml_cross_validator(
  sc,
  estimator = pipeline,
  estimator_param_maps = list(
    linear_regression = list(
      elastic_net_param = c(0, 0.25, 0.5, 0.75, 1), #c(0, 0.25, 0.75, 1)
      reg_param = seq(from = 0, to = 0.001, length.out = 50)#seq(from = 0.001, to = 1, length.out = 101)
    )
  ),
  evaluator = ml_regression_evaluator(
    sc,
    label_col = "resale_price_lg"
  ),
  num_folds = 10,
  parallelism = 10,
  seed = 1337
)
cv_model <- ml_fit(cv, split_train)
cv_model

cv_mod_eval <- ml_validation_metrics(cv_model)
cv_mod_eval |> arrange(rmse)
cv_model$best_model

```

#Saving pipeline and model
```{r}
ml_save(pipeline, path = "spark_pipeline", overwrite = TRUE)
ml_save(cv_model$best_model, path = "hdb_cv_model", overwrite = TRUE)
```
```{r}
library(plumber)

plumb(file = "hdb_plumber.R") |>
  pr_run(port = 8000)

```

#Disconnect From Spark
```{r}
spark_disconnect(sc)
```
```{r}
toString(12)
```

